---
title: "Hierarchical Successive Removal"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(hbayeseco)
library(rjags)
library(tidyverse)
library(bayesplot)
library(GGally)
library(ggExtra)
```




The model rely on successive discrete (Binomial) distribution which the size parameter can get very low. This renders Normal approximation unreliable. Using `Stan` with Hamiltonian Monte Carlo would require to marginalize over the successive Binomial distribution. This requires some burdensome calculus. Hence we will use `jags` with a standard MCMC.

## Fitting the simple independent model with successive removal technique

```{r data}
#dataList <- source(file = "Chap_10_2_Hierarchical_SuccRem2_Nivelle/data_Hierarchical_SuccRem2_Nivelle.txt")
data("SucRemNivelle")
```

```{r indep_model}
model_string <- "model
{

for (i in 1:I)
{

# Prior
# -----

# Prior on capture efficiency
	pi[i] ~ dbeta(1, 1)

# Prior on fish density
	d[i] ~ dgamma(1, 1)

# Number of fish on each site i (surface S[i] in the data set)
	Enu[i] <- S[i] * d[i]
	nu[i] ~ dpois(Enu[i]) T(,500)

# Likelihood
# ----------

	C1[i] ~ dbin(pi[i], nu[i])
	nur1[i] <- nu[i] - C1[i]
	C2[i] ~ dbin(pi[i], nur1[i])
	nur2[i] <- nur1[i] - C2[i]
	C3[i] ~ dbin(pi[i], nur2[i])

}

}"
```


```{r fit-ind}
model <- jags.model(file = textConnection(model_string), 
                    data = SucRemNivelle,
                    n.chains = 5)

# Inferences
update(model, n.iter = 1000)
posterior_sample <- coda.samples(model,
                       variable.names = c("pi", "d", "nu"),
                       n.iter = 10000,
                       thin = 10)
```

```{r trace-ind}
mcmc_trace(posterior_sample, pars = c("d[5]", "pi[5]"))
```

```{r df-ind}
df_ind <- extract_wider(posterior_sample)
```

```{r}
p <- df_ind %>% 
  select(`pi.5`, `d.5`) %>% 
  ggplot(aes(`d.5`, `pi.5`)) +
  geom_point(col = "red") + 
  geom_density_2d(col = "black") +
  theme_bw(base_size = 15L, base_line_size = 0) +
  ylim(c(0,1)) +
  theme(legend.position="none")+
  labs(x=expression(delta),
       y=expression(pi),title="Fish density vs Capture efficiency at site 5") 
ggExtra::ggMarginal(p, type = "densigram", bins = 40)
  
```

## Hierarchical modeling
```{r hier-model}
model_string_hier <- "model
{

# Priors
# ------

# Hyperprior for the trapping efficiency
mu.p ~ dbeta(1,1) 
log.u.p ~ dunif(0,10)
u.p <- exp(log.u.p)

a <- mu.p * u.p
b <- (1-mu.p) * u.p

# Compute standard deviation of p
var.p <- mu.p*(1-mu.p)*(1/(u.p+1))
sd.p <- sqrt(var.p)

# Hyperpriror for the density
mu.d ~ dgamma(1,1)
sd.d ~ dgamma(1,0.001)

# Parameterization of the Gamma distribution
c1 <- pow(mu.d,2) / pow(sd.d,2)
d1 <- mu.d / pow(sd.d,2)

for (i in 1:I)
{

# Prior on capture efficiency
	pi[i] ~ dbeta(a,b)

# Prior on fish density
	d[i] ~ dgamma(c1,d1)

# Number of fish (S[i] in the data set)
	Enu[i] <- S[i] * d[i]
	nu[i] ~ dpois(Enu[i]) T(,500)

}

# Likelihood
# ----------

for (i in 1:I)
{
	C1[i] ~ dbin(pi[i], nu[i])
	nur1[i] <- nu[i] - C1[i]
	C2[i] ~ dbin(pi[i], nur1[i])
	nur2[i] <- nur1[i] - C2[i]
	C3[i] ~ dbin(pi[i], nur2[i])
}

}
"

```

```{r fit-hier}
model_hier <- jags.model(
  file = textConnection(model_string_hier), 
  data = SucRemNivelle,
  n.chains = 5)

# Inferences
update(model_hier, n.iter = 2000)
posterior_sample_hier <- coda.samples(
  model_hier,
  variable.names = c("a", "b", "c1", "d1", "mu.d", "mu.p", 
                     "pi", "d", "nu"),
  n.iter = 20000,
  thin = 10)
```

```{r trace-hier}
mcmc_trace(posterior_sample_hier, 
           pars = c("a", "b", "c1", "d1", "mu.d", "mu.p","d[5]", "pi[5]"))
```

```{r}
df_hier <- extract_wider(posterior_sample_hier)

```

```{r}
phier <- df_hier %>% 
  select(`pi.5`, `d.5`) %>% 
  ggplot(aes(`d.5`, `pi.5`)) +
  geom_point(col = "red") + 
  geom_density_2d(col = "black") +
    ylim(c(0,1)) + xlim(c(0,0.3)) +
  theme_bw(base_size = 15L, base_line_size = 0) +
  theme(legend.position="none")+
  labs(x=expression(delta),
       y=expression(pi),title="Fish density vs Capture efficiency at site 5 for the hierarchical model") 
ggExtra::ggMarginal(phier, type = "densigram", bins = 40)
  
```

```{r}
df_hier %>% 
  select("a", "b", "c1", "d1", "mu_d", "mu_p") %>% 
  GGally::ggpairs(
  columnLabels = c("a", "b", "c", "d", "mu[delta]", "mu[pi]"),
  labeller = "label_parsed"
  )
```


## Fast sampling technique  with abundance index



```{r model-AI}
model_string_AI <- "model
{

# Priors
# ------

# Hyperprior for the trapping efficiency
mu.p ~ dbeta(1,1) 
log.u.p ~ dunif(0,10)
u.p <- exp(log.u.p)

a <- mu.p * u.p
b <- (1-mu.p) * u.p

# Compute standard deviation of p
var.p <- mu.p*(1-mu.p)*(1/(u.p+1))
sd.p <- sqrt(var.p)

# Hyperpriror for the density
mu.d ~ dgamma(1,1)
sd.d ~ dgamma(1,0.001)

# Parameterization of the Gamma distribution
c1 <- pow(mu.d,2) / pow(sd.d,2)
d1 <- mu.d / pow(sd.d,2)

for (i in 1:I)
{

# Prior on capture efficiency
	pi[i] ~ dbeta(a,b)

# Prior on fish density
	d[i] ~ dgamma(c1,d1)

# Number of fish (S[i] in the data set)
	Enu[i] <- S[i] * d[i]
	nu[i] ~ dpois(Enu[i]) T(,500)

}

# Hyperpriors for the mean abundance index
k ~ dgamma(1, 0.001)
phi ~ dbeta(1, 1)
f <- phi/(1-phi)

# Priors for the Abundance index
for (i in 1:I) {
  alpha.AI[i] <- k*f*d[i]
  mu.AI[i] ~ dgamma(alpha.AI[i], f)

}


# Likelihood
# ----------

for (i in 1:I) {
	C1[i] ~ dbin(pi[i], nu[i])
	nur1[i] <- nu[i] - C1[i]
	C2[i] ~ dbin(pi[i], nur1[i])
	nur2[i] <- nur1[i] - C2[i]
	C3[i] ~ dbin(pi[i], nur2[i])
  AI[i] ~ dpois(mu.AI[i])
  
}

# Predictive distribution of the density conditionnally upon
# observed abundance indices

for (i in 1:I_pred) {
  AI_pred[i] ~ dpois(mu.AI_pred[i])
  mu.AI_pred[i] ~ dgamma(alpha.AI_pred[i], f)
  alpha.AI_pred[i] <- k * d_pred[i] * f
  d_pred[i] ~ dgamma(c1,d1)
  # Link with observed abundance indices
}

}
"

```

We have the following abundance index data

```{r data-AI}
# Abundance indices data
# Nivelle

data_AI <- list(

AI =  c(
6, 17, 12, 17, 11, 22, 16, 25, 25, 15, 11,
4, 10, 6, 9, 24, 13, 16, 14, 10, 25, 4,
NA, 24, NA, 63, NA, NA, 21, 41, 44, NA, 5),

I_pred=65, 

AI_pred = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25,
26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48,
49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65)

)
```

```{r fit-AI}
model_AI <- jags.model(file = textConnection(model_string_AI), 
                    data =SucRemNivelle,#c(dataList$value, data_AI),
                    n.chains = 5)

# Inferences
update(model_AI, n.iter = 2000)
posterior_sample_AI <- coda.samples(
  model_AI,
  variable.names = c("a", "b", "c1", "d1", "mu.d", "mu.p", "sd.d", "sd.p",
                     "pi", "d", "nu", "phi", "k", "f", "mu.AI", 
                     "d_pred", "mu.AI_pred"),
  n.iter = 20000,
  thin = 10)
```

## Plotting the results

Figure 10.13 page 253, posterior distributions of parameters for the full AI + successive removal model

```{r plot-10-13}
df_AI <- extract_wider(posterior_sample_AI)
df_AI %>% 
  select("mu_d", "sd_d", "mu_p",  "sd_p", "phi", "k") %>% 
  GGally::ggpairs(
  columnLabels = c("mu[delta]", "sigma[delta]", "mu[pi]",  "sigma[pi]", "phi", "k"),
  labeller = "label_parsed",
  lower = list(continuous = wrap("points", alpha = 0.3,size = .5))
  ) + 
  theme_linedraw(base_size = 15L, base_line_size = 0) +
  theme(axis.ticks = element_line(size = 15/22))
```

Figure 10.14 page 249, marginal posterior distributions of salmon densities ($\delta$) for each site index.

```{r long-df}
fit_hier_l <- extract_longer(posterior_sample_hier) %>% mutate(model = "hierarchical")
fit_ind_l <- extract_longer(posterior_sample) %>%  mutate(model = "independent")
fit_ai_l <- extract_longer(posterior_sample_AI) %>%  mutate(model = "abundance_index")
```


```{r plot-10-14}
bind_rows(fit_hier_l, fit_ind_l, fit_ai_l) %>% 
  filter(str_detect(parameter, "^d\\.")) %>%
  rename("delta" = value) %>% 
  mutate(Site = as_factor(str_sub(parameter, 3))) %>%  # 6 = length("d.") + 1
  ggplot() +
  aes(x = Site, y = delta, fill = model ) +
  geom_boxplot(outlier.shape = NA, outlier.size = .25) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ylim(c(0,0.75)) +
  scale_fill_manual(values = c("gray10", "gray50", "gray90")) +
  ylab("Fish density") +
  theme_classic(base_size = 15L)
```

Figure 10.17 page 251: Posterior means of Salmon juvenile density (under hierarchical bayesian model) versus observed Abundance Indices.
```{r plot-10-17}
df_delta_AI  <- fit_hier_l %>% 
  filter(str_detect(parameter, "^d\\.")) %>%
  rename("delta" = value) %>% 
  mutate(Site = as_factor(str_sub(parameter, 3))) %>% 
  group_by(Site) %>% 
  summarise(mean_delta = mean(delta)) %>% 
  mutate(abundance = data_AI$AI)
df_delta_AI %>% 
  ggplot(aes(x = abundance, y = mean_delta)) +
  ylim(c(0, .4)) + xlim(c(0, 70)) +
  geom_point() +
  xlab("Abundance indices") +
  ylab("Estimated fish density") +
  theme_classic(base_size = 15L)
```

Figure 10.20 page 254: Predictive posterior distribution of Salmon juvenile density versus Abundance Indices (50% and 95% credible interval).

Warning: La légende du libre n'est pas correct, il s'agit d'intervalle de crédibilité à 80%.

```{r}
fit_ai_l %>% 
  filter(str_detect(parameter, "^d_pred.")) %>%
  mutate(AI = as.numeric(str_sub(parameter, 8))) %>% 
  rename("delta_pred" = value) %>% 
  group_by(AI) %>% 
  summarise(qs = quantile(delta_pred, 
                          c(.1, .25, .5, .75, .9)),
                          prob = c(.025, .25, .5, .75, .975)) %>% 
  pivot_wider(names_from = prob, values_from =  qs) %>% 
  ggplot(aes(x = AI, y = `0.5`)) +
  geom_ribbon(aes(ymin = `0.025`, ymax = `0.975`), fill = "gray50") +
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), fill = "gray75") +
  geom_line() +
  geom_point(data = df_delta_AI,
  mapping = aes(x = abundance, y = mean_delta)) +
  xlab("Abundance indices") +
  ylab("Posterior predictive of fish density") +
  theme_classic(base_size = 15L)

```

```{r}
fit_ai_l %>% 
  filter(str_detect(parameter, "^d_pred.")) %>%
  mutate(AI = as.numeric(str_sub(parameter, 8))) %>% 
  rename("delta_pred" = value) %>% 
  group_by(AI) %>% 
  summarise(qs = quantile(delta_pred, 
                          c(.025, .25, .5, .75, .975)),
                          prob = c(.025, .25, .5, .75, .975)) %>% 
  pivot_wider(names_from = prob, values_from =  qs) %>% 
  ggplot(aes(x = AI, y = `0.5`)) +
  geom_ribbon(aes(ymin = `0.025`, ymax = `0.975`), fill = "gray50") +
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), fill = "gray75") +
  geom_line() +
  geom_point(data = df_delta_AI,
  mapping = aes(x = abundance, y = mean_delta)) +
  xlab("Abundance indices") +
  ylab("Posterior predictive of fish density") +
  theme_classic(base_size = 15L)

```
