---
title: "Hierarchical Stock Recruitment"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.retina = 2
)
```

```{r setup}
library(hbm4ecology)
library(rstan)
library(tidyverse)
```


Parent E. and Rivot E. - An introduction to Hierarchcial Modeling of Ecological Data Chapman et al. 2012

Stan program for Hierarchical Stock-recruitment analysis (with covariates Latitude) Chapter 9 - section 9.3



# Model without covariates

Ricker model with lognormal errors and management related parametrization

- $\log(R_{k,t}) = h_{k}^{*} + \log(\frac{S_{k,t}}{1-h^{*}{t}}) - \frac{h^{*}_{k}}{S^{*}_{k}}S_{k,t} + \epsilon_{k,t}$
- $\epsilon_{k,t} \overset{iid}{\sim} Normal(0, \sigma_k^2)$

### Defining priors for the model 

- $h_k^* \sim Beta(1,1)$
- $\tau = \sigma^{-2} \sim Gamma(p = 10^{-3},q = 10^{-3})$ with $\sigma_k = k$, $\forall k$

With two different elicitation for $S^*_k$:

A more refined gamma elicitation:

- $\mu_{S^*} = 40 \text{eggs/$m^2$}$
- $CV_{S^*} = 1$
- $a = CV_{S^*}^{-2}$
- $b = \mu_{S^*}^{-1}CV_{S^*}^{-2}$
- $S_k^* \sim Gamma(a,b) \mathbf{1}_{S_k^* < 200}$

A simple uniform:

- $S_k^* \sim Uniform(0,200)$

### Writing the model in stan

```{r}
hsr_model_stan <- "
data {
  int n_riv; // Number of rivers
  int n_obs[n_riv]; // Number of observations by rivers
  int n; // Total number of observations
  int riv[n]; // River membership (indices k)
  vector[n] S; // Stock data
  vector[n] R; // Recruitment data
}

parameters {
  vector<lower=0, upper=200>[n_riv] Sopt;
  vector<lower=0, upper=1>[n_riv] hopt;
  real<lower=0> tau;
}

transformed parameters {
  vector<lower=0>[n_riv] alpha;
  vector<lower=0>[n_riv] beta ;
  real<lower=0> sigma;
  // real Copt;
  // real Ropt;
  vector[n] LogR;
  // real Slope;
  
  for (k in 1:n_riv) {
    alpha[k] = exp(hopt[k])/(1-hopt[k]) ;  
    beta[k] = hopt[k]/Sopt[k]     ;    
  }
  sigma = tau^-0.5        ;  
  // Copt = hopt*Sopt/(1-hopt);
  // Ropt = Sopt + Copt;
  // Slope = exp(alpha);
  for (t in 1:n) {
    LogR[t] = log(S[t]) + log(alpha[riv[t]]) - beta[riv[t]]*S[t] ;  // LogR[t] is the logarithm of the Ricker function  
  }
}

model {
  tau ~ gamma(.001, .001) ; // Diffuse prior for the variance
  for (k in 1:n_riv) {
    hopt[k] ~ beta(1,1) ;   
    Sopt[k] ~ uniform(0, 200) ;
  }
// 
    log(R) ~ normal(LogR, sigma) ;
}
"
```

```{r}
data("HsrAsalmon")
# data <- list(
# 
# n_riv = 13, 
# 
# n_obs = c(12,14,12,9,12,7,13,13,11,12,6,8,10),
# 
# n = 139, 
# 
# riv = c(
# 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 
# 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 
# 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 
# 4, 4, 4, 4, 4, 4, 4, 4, 4, 
# 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 
# 6, 6, 6, 6, 6, 6, 6, 
# 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 
# 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8, 
# 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 9, 
# 10, 10, 10, 10, 10, 10, 10, 10, 10, 10, 10, 10, 
# 11, 11, 11, 11, 11, 11, 
# 12, 12, 12, 12, 12, 12, 12, 12, 
# 13, 13, 13, 13, 13, 13, 13, 13, 13, 13),
# 
# S = c(0.957, 0.501, 2.286, 1.481, 1.596, 2.679, 2.323, 1.645, 2.210, 3.764, 2.470, 1.391, 
# 12.956, 12.084, 10.523, 2.762, 8.278, 6.943, 5.135, 1.540, 2.262, 3.437, 8.305, 5.231, 7.889, 1.621,
# 5.785, 6.504, 7.325, 10.052, 11.969, 8.660, 5.192, 2.354, 2.494, 3.172, 2.887, 2.921, 
# 2.399, 3.793, 0.989, 1.621, 1.538, 1.822, 3.872, 1.913, 2.364,
# 4.059, 7.377, 8.705, 6.032, 6.159, 7.653, 3.477, 7.250, 4.134, 6.441, 6.062, 6.013,
# 2.199, 1.844, 2.364, 1.395, 2.979, 2.388, 2.080, 
# 1.384, 3.584, 5.665, 4.057, 5.441, 1.254, 2.886, 3.513, 3.040, 3.548, 2.661, 1.727, 1.833, 
# 0.296, 1.284, 0.992, 0.366, 1.193, 1.496, 1.818, 1.258, 1.169, 2.813, 3.136, 1.189, 1.414, 
# 3.914, 16.788, 9.091, 1.678, 5.970, 15.327, 11.944, 7.955, 13.640, 16.008, 24.593, 
# 4.765, 2.467, 3.454, 9.036, 9.206, 3.540, 2.382, 4.680, 2.876, 1.021, 3.046, 5.428, 
# 11.381, 11.425, 18.308, 14.920, 17.584, 17.733, 
# 6.331, 5.269, 6.369, 4.913, 6.129, 5.782, 5.383, 2.376, 
# 37.010, 37.303, 28.337, 16.790, 37.955, 42.157, 32.309, 34.869, 24.022, 31.268), 
# 
# R = c(2.647, 0.444, 1.801, 0.573, 0.356, 0.390, 3.598, 2.357, 2.026, 0.665, 0.834, 1.433,
# 8.247, 1.328, 2.744, 4.511, 5.174, 0.846, 4.278, 1.237, 3.429, 4.491, 5.440, 2.797, 5.997, 0.442, 
# 13.278, 13.266, 9.062, 5.269, 2.919, 3.467, 4.080, 3.757, 3.828, 4.397, 4.640, 5.370, 
# 2.617, 3.141, 4.205, 3.655, 3.898, 3.333, 2.792, 3.877, 3.290, 
# 15.226, 27.525, 8.615, 16.634, 13.926, 16.805, 17.668, 25.368, 12.232, 18.723, 16.516, 21.778, 
# 5.114, 5.145, 4.207, 3.491, 3.412, 3.759, 4.248, 
# 17.341, 6.301, 11.328, 14.334, 15.589, 19.808, 12.307, 9.313, 6.717, 10.268, 18.339, 5.982, 5.817, 
# 3.727, 4.675, 5.681, 3.932, 3.653, 8.789, 9.800, 3.715, 4.419, 3.435, 2.991, 4.184, 2.309, 
# 18.657, 47.897, 37.326, 24.859, 42.626, 50.024, 42.996, 62.642, 76.854, 26.009, 14.744,
# 11.766, 3.594, 8.026, 6.119, 7.735, 7.152, 5.919, 9.419, 6.065, 4.016, 10.110, 4.438, 
# 20.154, 20.791, 26.458, 19.724, 16.823, 21.693, 
# 15.323, 13.202, 15.969, 17.781, 18.242, 16.253, 10.899, 5.724, 
# 150.747, 108.681, 104.854, 110.264, 111.579, 139.799, 124.208, 85.049, 73.261, 71.534)
# 
# ) 
```


```{r}
#options(mc.cores = parallel::detectCores())
# Number iteration for the "warm up" phase
n_warm <- 2500
# Number iteration for inferences
n_iter <- 5000
n_thin <- 1
# Number of chains
n_chains <- 4

#STAN_model_SRMargaree_Ricker_LogN_Management <- 
#  STAN_model_SRMargaree_Ricker_LogN_Management.stan


# ------------------------------------------------------------------------------------
# HMC config.
# ------------------------------------------------------------------------------------
# Run model
# ------------------------------------------------------------------------------------

# Load/check model, load data and inits

# model_code <- stanc(file = "./STAN_SRMargaree_Ricker_LogN_Management.stan")
model_name <-"HSR_Ricker_LogN_Management"
sm <- stan_model(model_code =  hsr_model_stan,
                 model_name =  model_name
)
params<-c("Sopt","hopt","tau")
# Inferences


fit <- sampling(object = sm,
                data = HsrAsalmon, 
                pars = NA, #params,
                chains = n_chains, 
                iter = n_iter, 
                warmup = n_warm, 
                thin = n_thin, 
                control = list("adapt_delta" = .85)
                #control = list("max_treedepth" = 12)
)

stan_rhat(fit,pars = params )
summary(fit)
# plot(fit, plotfun = "trace", pars = params, inc_warmup = TRUE)
# 
# plot(fit, plotfun = "hist", pars = params, include = TRUE)+ 
#   ggtitle("Posterior distributions for main parameters")
# pairs(x = fit, 
#       pars = params
#  )
# print(fit)
# plot(fit)
# traceplot(fit, "lp__", inc_warmup=TRUE)
# plot(fit, show_density = TRUE, ci_level = 0.5, fill_color = "purple")
# plot(fit, plotfun = "trace", pars = params, inc_warmup = TRUE)
# plot(fit, plotfun = "rhat") + ggtitle("Example of adding title to plot")
```

```{r}
ar <-  rstan::extract(fit, pars= "Sopt",  permuted=FALSE, inc_warmup=FALSE)
df <- lapply(seq(n_chains), 
              function(c) as_tibble(ar[,c,]) %>% mutate(iterations = seq.int(nrow(ar[,c,])),
                                                         chains = as_factor(eval(c)))) %>% 
  bind_rows()

df %>% pivot_longer(cols = starts_with("Sopt")) %>% 
  ggplot() +
  aes(y = log(value), group = name) +
  geom_boxplot()

```

```{r}
ar <-  rstan::extract(fit, pars= "hopt",  permuted=FALSE, inc_warmup=FALSE)
df <- lapply(seq(n_chains), 
              function(c) as_tibble(ar[,c,]) %>% mutate(iterations = seq.int(nrow(ar[,c,])),
                                                         chains = as_factor(eval(c)))) %>% 
  bind_rows()

df %>% pivot_longer(cols = starts_with("hopt"), values_to = "h*") %>% 
  ggplot() +
  aes(y = `h*`, group = name) +
  geom_boxplot()

```

#### With Gamma prior for S

```{r}
hsr_model_indep_S_gamma_stan <- "
data {
  int n_riv; // Number of rivers
  int n_obs[n_riv]; // Number of observations by rivers
  int n; // Total number of observations
  int riv[n]; // River membership (indices k)
  vector[n] S; // Stock data
  vector[n] R; // Recruitment data
}

parameters {
  vector<lower=0, upper=200>[n_riv] Sopt;
  vector<lower=0, upper=1>[n_riv] hopt;
  real<lower=0> tau;
}

transformed parameters {
  vector<lower=0>[n_riv] alpha;
  vector<lower=0>[n_riv] beta ;
  real<lower=0> sigma;
  // real Copt;
  // real Ropt;
  vector[n] LogR;
  // real Slope;
  
  for (k in 1:n_riv) {
    alpha[k] = exp(hopt[k])/(1-hopt[k]) ;  
    beta[k] = hopt[k]/Sopt[k]     ;    
  }
  sigma = tau^-0.5        ;  
  // Copt = hopt*Sopt/(1-hopt);
  // Ropt = Sopt + Copt;
  // Slope = exp(alpha);
  for (t in 1:n) {
    LogR[t] = log(S[t]) + log(alpha[riv[t]]) - beta[riv[t]]*S[t] ;  // LogR[t] is the logarithm of the Ricker function  
  }
}

model {
  tau ~ gamma(.001, .001) ; // Diffuse prior for the variance
  for (k in 1:n_riv) {
    hopt[k] ~ beta(1,1) ;   
    Sopt[k] ~ gamma(1, 1./1600) ;
  }
// 
    log(R) ~ normal(LogR, sigma) ;
}
"
```

```{r}
#options(mc.cores = parallel::detectCores())
# Number iteration for the "warm up" phase
n_warm <- 2500
# Number iteration for inferences
n_iter <- 5000
n_thin <- 1
# Number of chains
n_chains <- 4
model_name <-"HSR_Ricker_LogN_Management_S_gamma"
sm <- stan_model(model_code =  hsr_model_indep_S_gamma_stan,
                 model_name =  model_name
)
params<-c("Sopt","hopt","tau")
# Inferences


fit <- sampling(object = sm,
                data = HsrAsalmon, 
                pars = NA, #params,
                chains = n_chains, 
                iter = n_iter, 
                warmup = n_warm, 
                thin = n_thin,
                control = list("adapt_delta" = .85)#max_treedepth" = 12)
)

 stan_rhat(fit,pars = params )
# plot(fit, plotfun = "trace", pars = params, inc_warmup = TRUE)
# 
# plot(fit, plotfun = "hist", pars = params, include = TRUE)+ 
#   ggtitle("Posterior distributions for main parameters")
# pairs(x = fit, 
#       pars = params
# )
# print(fit)
# plot(fit)
# traceplot(fit, "lp__", inc_warmup=TRUE)
# plot(fit, show_density = TRUE, ci_level = 0.5, fill_color = "purple")
# plot(fit, plotfun = "trace", pars = params, inc_warmup = TRUE)
# plot(fit, plotfun = "rhat") + ggtitle("Example of adding title to plot")
```

```{r}

ar <-  rstan::extract(fit, pars= "Sopt",  permuted=FALSE, inc_warmup=FALSE)
df <- lapply(seq(n_chains), 
              function(c) as_tibble(ar[,c,]) %>% mutate(iterations = seq.int(nrow(ar[,c,])),
                                                         chains = as_factor(eval(c)))) %>% 
  bind_rows()

df %>% 
  pivot_longer(cols = starts_with("Sopt"), values_to = "Sopt") %>%
  ggplot() +
  aes(y = log(Sopt), group = name) +
  geom_boxplot(fill = "grey80")  

```

```{r}
ar <-  rstan::extract(fit, pars= "hopt",  permuted=FALSE, inc_warmup=FALSE)
df <- lapply(seq(n_chains), 
              function(c) as_tibble(ar[,c,]) %>% mutate(iterations = seq.int(nrow(ar[,c,])),
                                                         chains = as_factor(eval(c)))) %>% 
  bind_rows()

df %>% pivot_longer(cols = starts_with("hopt"), values_to = "h*") %>%
  ggplot() +
  aes(y = `h*`, group = name) +
  geom_boxplot() +
  geom_boxplot(fill = "grey80")  

```

### Hierarchical model with latitudinal covariates


```{r}
hsr_model_hier_S_gamma_stan <- "
data {
  int n_riv; // Number of rivers
  int n_obs[n_riv]; // Number of observations by rivers
  int n; // Total number of observations
  int riv[n]; // River membership (indices k)
  vector<lower=0>[n] S; // Stock data
  vector<lower=0>[n] R; // Recruitment data
  vector[n_riv] lat; //Latitude of rivers
  int n_pred;
  vector[n_pred] lat_pred;
}

transformed data {
  vector[n_riv] lat_rescaled;
//  real CV_S;
//  real a;
//  a = 1/(CV_S*CV_S);
  lat_rescaled = lat-mean(lat);
}

parameters {
  real<lower=0> tau;

  real<lower=0, upper=20> CV_S;
  real<lower=-5, upper=5> alpha;
  real<lower=-50, upper=50> beta;
  vector<lower=0, upper=200>[n_riv] Sopt;  
  
  
  real<lower=-5, upper=5> delta;
  real<lower=-50, upper=50> kappa;
  real<lower=0> tau_h;
  vector[n_riv] logit_h; // logit of hopt

}

transformed parameters {
  
  real<lower=0> sigma; // stdev for lognormal noise
  
//  vector[n_riv] log_mu_S; // log of the expected mean of Sopt
  vector<lower=0>[n_riv] mu_S; // Expected mean of Sopt  
  real<lower=0> a; // parameters of the gamma prior of Sopt
  vector<lower=0>[n_riv] b; // parameters of the gamma prior of Sopt
  
  real<lower=0> sigma_h; // stdev for logit of h 
  vector[n_riv] logit_mu_h; // logit of the expected mean of hopt
  vector<lower=0, upper=1>[n_riv] hopt;
  
  vector[n] LogR;
  
  

  sigma = tau^-0.5  ;
  sigma_h = tau_h^-0.5 ;

  for (k in 1:n_riv) {
    mu_S[k] = exp(alpha * (lat[k]) + beta) ;
//    mu_S[k] = exp(alpha * (lat_rescaled[k]) + beta) ;
//    mu_S[k] = exp(log_mu_S[k]);
    logit_mu_h[k] = delta * (lat[k]) + kappa ;
//    logit_mu_h[k] = delta * (lat_rescaled[k]) + kappa ;
  }
  
  a = 1/square(CV_S) ;
  for (k in 1:n_riv) {
    b[k] = 1/(mu_S[k]*square(CV_S)) ;
  }  
  hopt = inv_logit(logit_h) ;


  for (t in 1:n) {
    LogR[t] = hopt[riv[t]] + log(S[t]) - log(1 - hopt[riv[t]]) - hopt[riv[t]]/Sopt[riv[t]]*S[t] ;  // LogR[t] is the logarithm of the Ricker function  
  }
}

model {
//  alpha ~ uniform(-5, 5);
//  beta ~ uniform(-50, 50);
//  CV_S ~ uniform(0, 20);
  
//  delta ~ uniform(-5, 5);
//  kappa ~ uniform(-50, 50);


  tau ~ gamma(.001, .001) ; // Diffuse prior for the precision
  tau_h ~ gamma(.001, .001) ; // Diffuse prior for the precision



  for (k in 1:n_riv) {
    logit_h[k] ~ normal(logit_mu_h[k], sigma_h);
    Sopt[k] ~ gamma(a, b[k]) ; // T[0,200] ;
  }

  
  log(R) ~ normal(LogR, sigma) ;
}

generated quantities {
  real a_pred;
  vector[n_pred] b_pred;
  vector[n_pred] Sopt_pred;
  vector[n_pred] logit_hopt_pred;
  vector<lower=0,upper=1>[n_pred] hopt_pred;
  // Posterior Predictive for new rivers with latitude lat_pred[p]
  a_pred = 1/(CV_S*CV_S);
  for ( p in 1 : n_pred ) {
//    b_pred[p] = a/(exp(alpha * lat_pred[p] + beta - alpha * mean(lat)));
    b_pred[p] = a_pred*(1/exp(alpha * lat_pred[p] + beta));
    Sopt_pred[p] = gamma_rng(a_pred,b_pred[p]); //T[0,200];
//    logit_hopt_pred[p] = normal_rng(exp(delta * lat_pred[p] + kappa - delta * mean(lat)), sigma_h);`
    logit_hopt_pred[p] = normal_rng(delta * lat_pred[p] + kappa, sigma_h);
    hopt_pred[p] = inv_logit(logit_hopt_pred[p]);
  } // end loop on predictions p
} 
"
```

```{r}
data_lat <- list(
  lat = c(43.5, 48.5, 50.5, 53, 53.98515, 54.5, 55, 55, 55, 57, 57, 61, 64),
  lat_pred = c(46, 52, 59, 63), 
  n_pred = 4
)

data_hier <- c(HsrAsalmon, data_lat)
```

```{r}
model_name_hier <-"HSR_Ricker_LogN_Management_S_gamma_hier"
sm_hier <- stan_model(model_code =  hsr_model_hier_S_gamma_stan,
                 model_name =  model_name_hier)
params<-c("Sopt","hopt","tau")
# Inferences


fit_hier <- sampling(object = sm_hier,
                data = HSrAsalmon, 
                pars = NA, #params,
                chains = n_chains,
                # init = lapply(seq(4), function(x)
                #   list(alpha = 0, beta = 0)),
                iter = 5000, 
                warmup = 2500, 
                thin = n_thin,
                control = list("adapt_delta"= .85)#list("max_treedepth" = 12)
)

stan_rhat(fit_hier,pars = params )
traceplot(fit_hier, "lp__", inc_warmup=FALSE)
pairs(fit_hier, pars = c("Sopt[1]", "hopt[1]"))
# plot(fit_hier, plotfun = "trace", pars = params, inc_warmup = TRUE)
# 
# plot(fit_hier, plotfun = "hist", pars = params, include = TRUE)+ 
#   ggtitle("Posterior distributions for main parameters")
# pairs(x = fit_hier, 
#       pars = params
# )
# print(fit_hier)
# plot(fit_hier)

# plot(fit_hier, show_density = TRUE, ci_level = 0.5, fill_color = "purple")
# plot(fit_hier, plotfun = "trace", pars = params, inc_warmup = TRUE)
# plot(fit_hier, plotfun = "rhat") + ggtitle("Example of adding title to plot")
```

### Results 

```{r}
names_ind <- c("Nivelle","Oir","Frome","Dee","Burrishoole","Lune","Bush",
"Mourne","Faughan","Girnock Burn","North Esk","Laerdalselva","Ellidaar")

names_pred <- c("Nivelle","new 45?","Oir","new 50?","Frome","Dee","Burrishoole","Lune","new 55?","Bush",
"Mourne","Faughan","Girnock Burn","North Esk","new 60?","Laerdalselva","Ellidaar","new 65?")
```


Plotting Figure 9., Page 210

```{r}
Sopt_hier <- fit_hier %>% extract_longer() %>% 
  filter(str_detect(parameter, "Sopt\\.")) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>% 
  group_by(Site) %>% summarise(Sopt_med = median(value))%>% 
  select(Sopt_med) %>% as_vector()
```


```{r}
hopt_hier <- fit_hier %>% extract_longer() %>% 
  filter(str_detect(parameter, "hopt\\.")) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>% 
  group_by(Site) %>% summarise(hopt_med = median(value))%>% 
  select(hopt_med) %>% as_vector()
```


```{r}
Sopt_ind <- fit %>% extract_longer() %>% 
  filter(str_detect(parameter, "Sopt\\.")) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>% 
  group_by(Site) %>% summarise(Sopt_med = median(value)) %>% 
  select(Sopt_med) %>% as_vector()
```


```{r}
hopt_ind <- fit %>% extract_longer() %>% 
  filter(str_detect(parameter, "hopt\\.")) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>% 
  group_by(Site) %>% summarise(hopt_med = median(value)) %>% 
  select(hopt_med) %>% as_vector()
```


```{r}
max_S <- sapply(seq_along(Sopt_ind), function(k) max(HsrAsalmon$S[HsrAsalmon$riv == k]))
Splot <- seq(0,100,length.out=1000)
Rplot_ind <- lapply(seq_along(Sopt_ind),
  function(k) {
    tibble(S = Splot[1:ceiling(15*max_S[k])],
           river = as_factor(names_ind[k]),
           R = exp( log(Splot[1:ceiling(15*max_S[k])]) + hopt_ind[k] - 
                     log(1-hopt_ind[k]) - (hopt_ind[k]/Sopt_ind[k])*Splot[1:ceiling(15*max_S[k])] ))
    })%>% 
  bind_rows()
```


```{r}
Rplot_hier <- lapply(seq_along(Sopt_hier),
  function(k) {
    tibble(S = Splot[1:ceiling(15*max_S[k])],
           river = as_factor(names_ind[k]),
           R = exp( log(Splot[1:ceiling(15*max_S[k])]) + hopt_hier[k] - 
                     log(1-hopt_hier[k]) - (hopt_hier[k]/Sopt_hier[k])*Splot[1:ceiling(15*max_S[k])] ))
    })%>% 
  bind_rows()



# df_Rplot_ind <- Rplot_ind %>% as_tibble() %>% 
#   mutate(S = Splot) %>% 
#   pivot_longer(cols = -S, names_to = "river", values_to = "R") %>%
#   mutate(river = as.numeric(str_sub(river, 2))) %>% 
#   mutate(river = factor(names_ind[river], levels = names_ind ))
# 
# df_Rplot_hier <- Rplot_hier %>% as_tibble() %>% 
#   mutate(S = Splot) %>% 
#   pivot_longer(cols = -S, names_to = "river", values_to = "R") %>%
#   mutate(river = as.numeric(str_sub(river, 2))) %>% 
#   mutate(river = factor(names_ind[river], levels = names_ind )) 

  

# df_Rplot_hier %>% group_by(river) %>% 
#   mutate(max_S = data$max_S)
```


```{r}
tibble(R = HsrAsalmon$R, S = HsrAsalmon$S, 
       river = factor(names_ind[HsrAsalmon$riv], levels = names_ind)) %>% 
  ggplot() +
  aes(x = S, y = R) +
  geom_point() +
  facet_wrap(~ river, scales = "free") +
  geom_line(data = Rplot_hier, size = 1) + 
  geom_line(data = Rplot_ind, size = 1, linetype = "dashed") +
  theme_classic(base_size = 15L) #+
#  scale_x_continuous(limits = )
```


```{r}
fit_hier_l <- extract_longer(fit_hier) %>% mutate(model = "hierarchical")
fit_ind_l <- extract_longer(fit) %>%  mutate(model = "independent")

fit_l <- bind_rows(fit_hier_l, fit_ind_l)

fit_l %>% 
  filter(str_detect(parameter, "Sopt\\.")) %>%
  rename("S*" = value) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>%  # 6 = length("Sopt.") + 1
  ggplot() +
  aes(x = Site, y = log(`S*`), fill = model ) +
  geom_boxplot(outlier.shape = NA, outlier.size = .25) +
  scale_x_discrete(labels = names_ind, guide = guide_axis(angle = 90)) +
  ylim(c(-3,6)) +
  scale_fill_manual(values = c("gray50", "gray90")) +
  ylab("log(S*)") +
  theme_classic(base_size = 15L)
```


```{r}
fit_l %>% 
  filter(str_detect(parameter, "hopt\\.")) %>%
  rename("h*" = value) %>% 
  mutate(Site = as_factor(str_sub(parameter, 6))) %>% 
  ggplot() +
  aes(x = Site, fill = model, y = `h*`) +
  geom_boxplot(outlier.shape = NA,outlier.size = .25) +
  scale_x_discrete(labels = names_ind, guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("gray50", "gray90"))+
  theme_classic(base_size = 15L)
```


Plotting marginal posterior probability shapes of the parameters $\alpha$, $\beta$, $\delta$, $\kappa$, $\sigma$, $\tau_{h*}$ (Fig. 9.12 page 216)

```{r}
fit_hier_l %>% 
  filter(parameter  %in% c("alpha", "beta", "delta", "kappa", "sigma", "sigma_h")) %>% 
  ggplot() + aes(x = value) +
  geom_density(adjust = 2) +
  facet_wrap(~ parameter, ncol = 2, scales = "free") +
  theme_classic(base_size = 15L) 
```

Marginal posterior distributions and posterior predictive of $\log(S*)$ for the hierarhical model (fig. 9.13 page 217)

```{r}
reg_param <- fit_hier %>% extract_wider() %>% select(alpha, beta) %>% 
  summarise(med_a = median(alpha), med_b = median(beta))

fit_data <- fit_hier_l %>% 
  filter(str_detect(parameter, "Sopt\\.")) %>%
  rename("S*" = value) %>% 
  mutate(id = as.factor(str_sub(parameter, 6)),
    Latitude = data_hier$lat[as.numeric(str_sub(parameter, 6))],
         type = "data") 

fit_pred <- fit_hier_l %>% 
  filter(str_detect(parameter, "Sopt_")) %>%
  rename("S*" = value) %>% 
  mutate(id = as.factor(str_sub(parameter, 6)),
    Latitude = data_hier$lat_pred[as.numeric(str_sub(parameter, 11))],
         type = "pred") 

fit_data %>% bind_rows(fit_pred) %>% 
  ggplot() +
  aes(x = Latitude, group = id,  y = log(`S*`), fill = type) +
  geom_boxplot(outlier.shape = NA,outlier.size = .25, varwidth = FALSE) +
#  scale_x_discrete(labels = names_ind, guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("gray90", "gray50")) +
  geom_abline(slope = reg_param$med_a, intercept = reg_param$med_b) + #-reg_param$med_a*mean(data_hier$lat)) +
  ylim(c(-2,5)) +
  ylab("log(S*)") +
  theme_classic(base_size = 15L)
```

Marginal posterior distributions and posterior predictive of $h*$ for the hierarchical model (fig. 9.14 page 218)

```{r}
reg_param <- fit_hier %>% extract_wider() %>% select(delta, kappa) %>% 
  summarise(med_d = median(delta), med_k = median(kappa))

fit_data <- fit_hier_l %>% 
  filter(str_detect(parameter, "^hopt\\.")) %>% # detect hopt. without logit_hopt.
  rename("h*" = value) %>% 
  mutate(id = as.factor(str_sub(parameter, 6)),
    Latitude = data_hier$lat[as.numeric(str_sub(parameter, 6))],
         type = "data") 

fit_pred <- fit_hier_l %>% 
  filter(str_detect(parameter, "^hopt_")) %>%
  rename("h*" = value) %>% 
  mutate(id = as.factor(str_sub(parameter, 6)),
    Latitude = data_hier$lat_pred[as.numeric(str_sub(parameter, 11))],
         type = "pred") 

fit_data %>% bind_rows(fit_pred) %>% 
  ggplot() +
  aes(x = Latitude, group = id,  y = `h*`, fill = type) +
  geom_boxplot(outlier.shape = NA,outlier.size = .25, varwidth = FALSE) +
#  scale_x_discrete(labels = names_ind, guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("gray90", "gray50")) +
  annotate(x = seq(42, 65, .1), 
           y = 1/(1+exp(-reg_param$med_d*seq(42, 65, .1) -reg_param$med_k)),
           geom = "line") +
  # geom_abline(slope = reg_param$med_d, 
  #             intercept = reg_param$med_k) + # -reg_param$med_d*mean(data_hier$lat)) +
  ylim(c(0,1)) +
#  ylab("log(S*)") +
  theme_classic(base_size = 15L, base_rect_size = 1)
```
