---
title: "Biomass Production Model: The Namibian Hake Fishery"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hbayeseco)
library(rstan)
library(tidyverse)
library(GGally)
library(posterior)
library(bayesplot)
```

## State-space modeling of a biomass production model

#### Process equation for the underlying dynamics of the biomass

- $\frac{\mbox{d}B_t}{\mbox{d}t} = h(B_t)$
- $h(B_t) = r\times B_t \times (1-\frac{B_t}{K})$ the production function
- $r \sim Uniform(0.01, 3)$ is the population intrinsic growth
- $K \sim Uniform(100, 15000)$ the carrying capacity

$h(B_t)$ is high when $B_t \approx 0$ and is null when $B_t = K$

Introducing the observed harvest between time $t$ and $t+1$,  $c_t$ and a log normal random noise:

- $B_{t+1} = (B_t + h(B_t) - c_t)\times e^{\epsilon_{t+1}}$
- $\epsilon_{t+1} \sim Normal(0, \sigma^2_p)$, the environmental noise
- $\sigma^2_p = \sigma^2$, with $\log(\sigma^2) \sim Uniform(-20,20)$
- with $B_1 = K e^{\epsilon_1}$

#### Observation equation to link the data to the hidden process

- $i_t = q \times B_t \times e^{\omega_t}$ the relative abundance index
- $q$ such that $log(q) \sim Uniform (-20, 20)$ the catchability parameter
- $\omega_t \sim Normal(0, \sigma_o^2)$ the observation noise
- $\sigma^2_o = \sigma^2$, with $\log(\sigma^2) \sim Uniform(-20,20)$

#### In state-space modeling notation:

\[
\begin{cases}
[Z_{t+1} | Z_t, \theta_1] \\
[y_t | Z_t, \theta_2]
\end{cases}
\]
where

- $Z_t = B_t$
- $y_t = i_t$
- $\theta_1 = \{r, K, \sigma_p \}$
- $\theta_2 = \{q, \sigma_o \}$


### Writing the model in Stan

```{r stan-model}
stan_model_biomprod <- "
data {
  int N;
  int Year[N];
  vector<lower=0>[N] C;
  vector<lower=0>[N-1] I;
  int NF;
}

parameters {
  real<lower=0> r;
  real<lower=100, upper=15000> K;
  real<lower=-20, upper=20> log_q;
  real<lower=-20, upper=20> log_sig2;
  vector<lower=0>[N] P; //B_t/K
}

transformed parameters {
  real q;
  real sig;
  vector<lower=0>[N] B;
  
  q = exp(log_q);
  sig = sqrt(exp(log_sig2));
  B = K*P;
}

model {
  vector[N-1] h_P;

  P[1] ~ lognormal(0, sig);
  for (t in 1:(N-1)) {
      h_P[t] = r * K * P[t] * (1 - P[t]);
      P[t+1] ~ lognormal(log(P[t] + h_P[t]/K - C[t]/K), sig);
      I[t] ~ lognormal(log(q*K*P[t+1]), sig);
  }
}

generated quantities {
  // Abundance prediction
  vector[N] I_pred;
  vector[NF] B_pred100;
  vector[NF] B_pred500;
  
  for (t in 1:N) {
    I_pred[t] = lognormal_rng(log(q*K*P[t]), sig);
  }
  // Forecasting the biomass

  B_pred100[1] = (B[N] + r*B[N]*(1-B[N]/K) - 100) * lognormal_rng(0, sig);
  B_pred500[1] = (B[N] + r*B[N]*(1-B[N]/K) - 500)* lognormal_rng(0, sig);
  for (t in 1:(NF-1)) {
  B_pred100[t+1] = (B_pred100[t] + r*B_pred100[t]*(1-B_pred100[t]/K) - 100) * lognormal_rng(0, sig);
  B_pred500[t+1] = (B_pred500[t] + r*B_pred500[t]*(1-B_pred500[t]/K) - 500) * lognormal_rng(0, sig);
  }
} 

"
```

```{r data}
# dat <- read.table("Chap_11_1_SSM_BiomProd_NamibianHake/data_table.txt", 
#                   header = TRUE)
# dat_list <- list(N = nrow(dat),
#                  NF = 5,
#                  Year = dat$Year,
#                  C = dat$C,
#                  I = dat$I[2:length(dat$I)])
data(BioprodNamibianHake, package = "hbayeseco")
dat_list <- BioprodNamibianHake
dat_list$I <- BioprodNamibianHake$I[2:25]
dat_list$NF <- 5
```

```{r fit-model}
model_name <-"BiomProd_NamibianHake"
sm_hier <- stan_model(model_code =  stan_model_biomprod,
                 model_name =  model_name)

fit_hier <- sampling(object = sm_hier,
                data = dat_list, 
                pars = NA, #params,
                chains = 4,
                init = lapply(seq(4), function(x)
                   list(P = rep(.5, 25))),
                iter = 4000, 
                warmup = 2000, 
                thin = 1,
                control = list("adapt_delta" = .9)#"max_treedepth" = 12)
)
```


```{r diagnostic-plot}
pars = c("r", "K", "q", "sig")
pairs(fit_hier, pars = c("r", "K", "q", "sig") )
traceplot(fit_hier, pars)
stan_rhat(fit_hier)
mcmc_trace(fit_hier, pars = pars)
rhat(fit_hier, pars = pars)
```
```{r}
pairs(fit_hier, pars = pars)
```


## Analysis 

```{r analysis}
df_w <- extract_wider(fit_hier)
df_l <- extract_longer(fit_hier)

```

```{r}
df_w %>% 
  mutate(sig2 = sig * sig,
         B_MSY = K/2,
         C_MSY = r*K/4) %>% 
  select(r, K, q, sig2, B_MSY, C_MSY) %>% 
  pivot_longer(cols = everything(), names_to = "parameter") %>% 
  group_by(parameter) %>% 
#  filter(parameter %in% c("r", "K", "q", "sig")) %>% 
  summarise(Mean = mean(value),
            Sd = sd(value),
            `q.025` = quantile(value, .025),
            Median = median(value),
            `q.975` = quantile(value, .975)) %>% 
  knitr::kable(digits = 5)
```

- $B_{MSY} = \arg \max_{B} h(B) = K/2$
- $C_{MSY} = g(B_{MSY}) = rK/4$ 

```{r}
df_w %>% 
  mutate(C_MSY = r*K/4) %>% 
  ggpairs(c("r" , "K", "q", "C_MSY"))
```

```{r}
df_l %>% 
  filter(str_detect(parameter, "I_pred")) %>% 
  mutate(
    id = as.numeric(str_sub(parameter, str_length("I_pred.")+1)),
    Year = BioprodNamibianHake$Year[id])  %>% 
  ggplot(aes(x = Year, y = value, group = Year)) +
  geom_boxplot(outlier.shape = NA) +
  annotate("line", x = BioprodNamibianHake$Year, y = BioprodNamibianHake$I) +
#  geom_line(data = dat, mapping = aes(y = I)) +
  ylim(c(0,2)) +
  ylab("Abundance indices") +
  scale_x_continuous(breaks = BioprodNamibianHake$Year,
                     labels = BioprodNamibianHake$Year, 
                     guide = guide_axis(angle = 90)) +
  theme_classic(base_size = 15)
  
```

Forecasting the biomass for years 1989 to 1993
```{r}
df_100 <- df_l %>% 
  filter(str_detect(parameter, "B_pred100\\.")) %>% 
  mutate(
    id = as.numeric(str_sub(parameter, str_length("B_pred100.")+1)),
    Year = seq(1989, 1993)[id], 
    data_type = "Pred100")    

df_500 <- df_l %>% 
  filter(str_detect(parameter, "B_pred500\\.")) %>% 
  mutate(
    id = as.numeric(str_sub(parameter, str_length("B_pred500.")+1)),
    Year = seq(1989, 1993)[id], 
    data_type = "Pred500")    

df_l %>%
  filter(str_detect(parameter, "B\\.")) %>%
  mutate(
    id = as.numeric(str_sub(parameter, str_length("B.")+1)),
    Year = BioprodNamibianHake$Year[id],
    data_type = "data")  %>%
  bind_rows(df_100, df_500) %>% 
  ggplot(aes(x = as_factor(Year), y = value)) +
  geom_boxplot(aes(fill = as_factor(data_type)),outlier.shape = NA, show.legend = FALSE) +
  ylim(c(0,10000)) +
  ylab("Biomass (x 1000 tons)") +
  xlab("Years") +
  scale_x_discrete( guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("white", "gray90", "gray50"))+
  theme_classic(base_size = 15)
```


### Comparing Schaefer-type versus Fox-type prodcution function

Recall that Schaefer type production function is given by:
\[ h(B_t) = r \times B_t \times (1-\frac{B_t}{K}) \],
with its associated fisheries management points:
\begin{cases}
C_{MSY} = \frac{r \times K}{4} \\
B_{MSY} = \frac{K}{2}
\end{cases}

The value of the Fox model are as follows:
\[ $h(B_t) = r \times B_t \times (1-\frac{\log(B_t)}{\log(K)}) \],
with its associated fisheries management points:
\begin{cases}
C_{MSY} = \frac{r \times K \times e^{-1}}{\log(K)} \\
B_{MSY} = K \times e^{-1}
\end{cases}


```{r stan-model-fox}
stan_model_fox <- "
data {
  int N;
  int Year[N];
  vector<lower=0>[N] C;
  vector<lower=0>[N-1] I;
  int NF;
}

parameters {
  real<lower=0> r;
  real<lower=100, upper=15000> K;
  real<lower=-20, upper=20> log_q;
  real<lower=-20, upper=20> log_sig2;

  vector<lower=0>[N] P; //B_t/K
}

transformed parameters {
  real q;
  real sig;
  vector<lower=0>[N] B;
  
  q = exp(log_q);
  sig = sqrt(exp(log_sig2));
  B = K*P;
}

model {
  vector[N-1] h_P;

  P[1] ~ lognormal(0, sig);
  for (t in 1:(N-1)) {
      h_P[t] = r * K * P[t] * (1 - log(K*P[t])/log(K));
      P[t+1] ~ lognormal(log(P[t] + h_P[t]/K - C[t]/K), sig);
      I[t] ~ lognormal(log_q+log(K*P[t+1]), sig);
  }
}

generated quantities {
  // Abundance prediction
  vector[N] I_pred;
  vector[NF] B_pred100;
  vector[NF] B_pred500;
  
  for (t in 1:N) {
    I_pred[t] = lognormal_rng(log(q*B[t]), sig);
  }
  // Forecasting the biomass

  B_pred100[1] = (B[N] + r*B[N]*(1-log(B[N])/log(K)) - 100) * exp(normal_rng(0, sig));
  B_pred500[1] = (B[N] + r*B[N]*(1-log(B[N])/log(K)) - 500)* exp(normal_rng(0, sig));
  for (t in 1:(NF-1)) {
  B_pred100[t+1] = (B_pred100[t] + r*B_pred100[t]*(1-log(B_pred100[t])/log(K)) - 100) * exp(normal_rng(0, sig));
  B_pred500[t+1] = (B_pred500[t] + r*B_pred500[t]*(1-log(B_pred500[t])/log(K)) - 500) * exp(normal_rng(0, sig));
  }
} 

"
```


```{r fit-model-fox}
model_name <-"BiomProd_NamibianHake_fox"
sm_fox <- stan_model(model_code =  stan_model_fox,
                 model_name =  model_name)

fit_fox <- sampling(object = sm_fox,
                data = dat_list, 
                pars = NA, #params,
                chains = 4,
                 init = lapply(seq(4), function(x)
                   list(K = 5000, r = 0.3, P = rep(.5, 25))),
                iter = 5000, 
                warmup = 2500, 
                thin = 1#,
              #  control = list("max_treedepth" = 12)
)
```

```{r}
pars = c("q", "K", "r", "sig")
stan_rhat(fit_fox)
pairs(fit_fox, pars = pars)
```


```{r compare}
df_l_fox <- extract_longer(fit_fox) %>% 
  mutate(model = "Fox")

df_100_fox <- df_l_fox %>% 
  filter(str_detect(parameter, "B_pred100\\.")) %>% 
  mutate(
    id = as.numeric(str_sub(parameter, str_length("B_pred100.")+1)),
    Year = seq(1989, 1993)[id], 
    data_type = "Pred100")  

df_l %>% mutate(model = "Schaefer") %>% 
  bind_rows(df_l_fox) %>% 
    filter(str_detect(parameter, "B\\.")) %>%
  mutate(
    id = as.numeric(str_sub(parameter, str_length("B.")+1)),
    Year = BioprodNamibianHake$Year[id],
    data_type = "data")  %>%
  bind_rows(df_100 %>%  mutate(model = "Schaefer"), df_100_fox) %>% 
  ggplot(aes(x = as_factor(Year), y = value)) +
  geom_boxplot(aes(fill = model),outlier.shape = NA) +
  geom_vline(xintercept = 25.5, linetype = "dashed") +
  ylim(c(0,10000)) +
  ylab("Biomass (x 1000 tons)") + xlab("Years") +
  scale_x_discrete( guide = guide_axis(angle = 90)) +
  scale_fill_manual(values = c("gray90", "gray50"))+
  theme_classic(base_size = 15)
```

```{r}
df_w %>% mutate(C_MSY = r*K/4,
                Model = "Schaefer") %>% 
  bind_rows(extract_wider(fit_fox) %>% 
              mutate(C_MSY = r*K*exp(-1)/log(K),
                     Model = "Fox")) %>% 
  ggplot(aes(x = C_MSY, linetype = Model)) +
  geom_density(adjust = 2) +
  theme_classic(base_size = 15)  
```

```{r}
df_w %>% mutate(B_MSY = K/2,
                Model = "Schaefer") %>% 
  bind_rows(extract_wider(fit_fox) %>% 
              mutate(B_MSY = K*exp(-1),
                     Model = "Fox")) %>% 
  ggplot(aes(x = B_MSY, linetype = Model)) +
  geom_density(adjust = 2) +
  theme_classic(base_size = 15)  
```

We now look at the chance of the biomass level to be under the Maximum Sustainable Yields biomass. Under the Schaefer production model:

```{r}
df_w %>% mutate(B_MSY = K/2) %>% 
  mutate(Depletion_100 = B_pred100.5 < B_MSY,
         Depletion_500 = B_pred500.5 < B_MSY) %>% 
  summarise(`Depletion (100K scenario)` = mean(Depletion_100),
            `Depletion (500K scenario)` = mean(Depletion_500)) %>% 
  knitr::kable()
```

Under the Fox production model:

```{r}
fit_fox %>% extract_wider() %>% 
  mutate(B_MSY = K*exp(-1)) %>% 
  mutate(Depletion_100 = B_pred100.5 < B_MSY,
         Depletion_500 = B_pred500.5 < B_MSY) %>% 
  summarise(`Depletion (100K scenario)` = mean(Depletion_100, na.rm = TRUE),
            `Depletion (500K scenario)` = mean(Depletion_500, na.rm = TRUE)) %>% 
  knitr::kable()
```
